Rao-Blackwellized smoother for Conditionally Linear and Gaussian models

### 2016-11-15 TexArxiv/clgm.tex
MAJ abstract

\begin{abstract}
This paper focuses on Sequential Monte Carlo based approximations of smoothing distributions in conditionally linear and Gaussian state spaces. Only a few methods have been proposed to reduce Monte Carlo variance of smoothers with Rao Blackwellisation steps for these models. The algorithm proposed by  \cite{briers:doucet:maskell:2010} extends the Bryson-Frazier smoother for Gaussian linear state space models using the generalized two-filter formula with Rao Blackwellisation steps. Other Rao Blackwellised Sequential Monte Carlo smoothers introduced in \cite{lindsten:bunch:sarkka:schon:godsill:2015,sarkka:bunch:godsill:2012,lindsten:bunch:godsill:schon:2013} are based on the forward backward decomposition of the marginal smoothing distribution % In \cite{kim:1994}, a structural approximation of the model is used to simplify the backward pass while \cite{lindsten:bunch:sarkka:schon:godsill:2015,sarkka:bunch:godsill:2012,lindsten:bunch:godsill:schon:2013} use Rao Blackwellisation steps
 and use explicit marginalization of the linear states in both forward and backward passes. In this paper, we propose an overview of these algorithms and introduce an additional Rao Blackwellisation step for all procedures which allows to sample new particles in the backward pass to improve the performance of the algorithm.  The procedures are applied to commodity markets which are described using a two factor model based on the spot price and a convenience yield. Parameters are estimated using  an Expectation Maximization and the performance of the proposed procedure is assessed with Chicago Mercantile Exchange crude oil data.

>>>>

\begin{abstract}
This paper focuses on Sequential Monte Carlo based approximations of smoothing distributions in conditionally linear and Gaussian state spaces. To reduce Monte Carlo variance of smoothers, it is typical in these models to use Rao-Blackwellisation. Particle approximation is used to sample the sequence of the hidden regimes; the gaussian state sequence are then explicitly integrated conditional to the sequence of regimes and observations, using variants of the Kalman filter / smoother. The first successful attempt to use Rao-Blackwellisation for smoothing has been presented in  \cite{briers:doucet:maskell:2010} which extends the Bryson-Frazier smoother for Gaussian linear state space models using the generalized two-filter formula together with Kalman filter / smoother. More recently, in a series of article, \cite{lindsten:bunch:sarkka:schon:godsill:2015,sarkka:bunch:godsill:2012,lindsten:bunch:godsill:schon:2013} proposed to use a  forward backward decomposition of the smoothing distribution, mimicking the Rauch-Tung-Striebel smoother for the regime combined with backward Kalman updates.  % In \cite{kim:1994}, a structural approximation of the model is used to simplify the backward pass while \cite{lindsten:bunch:sarkka:schon:godsill:2015,sarkka:bunch:godsill:2012,lindsten:bunch:godsill:schon:2013} use Rao Blackwellisation steps
In this paper, we provide a survey of these algorithms in a common unifying framework. We then investigate the benefit of introducing an additional rejuvenation step  which allows to sample at each time instant new  regime conditional to the forward and backward particles.  To support our claims, a limited Monte Carlo study is presented. These procedures are applied to commodity markets which are described using a  factor model based on the spot price and a convenience yield for crude oil data.
